\section{Storing historical certificates for Ouroboros Genesis}\label{sec:storing historical certs}

Nodes bootstrapping via Ouroboros Genesis (to minimize trust assumptions) need to be able to assess the weight of competing historical chains in order to resist adversarial \emph{long-range attacks} \parencite{genesis-implementation-documentation}.
Computing the weight of a chain requires the corresponding certificates indicating which blocks are boosted.
See \cref{sec:weighted genesis} for more details.

Even if a syncing node did not encounter any adversarial challengers (and hence did not ever need to compute weight), it still must get all certificates:
Indeed, \emph{other} peers syncing from that node (e.g., in the future) might encounter adversaries and therefore require those certificates.
Briefly, there is a \enquote{duty to remember} certificates even if they are not directly useful for the caught-up node anymore.

We describe the protocol for retrieving historical certificates in \cref{sec:certificate-diffusion}, the associated storage component in \cref{sec:cert db}, and the adjustment to comparing chains by weight instead of density in \cref{sec:weight not length}.
Furthermore, it is necessary to prevent block synchronization from outpacing certificate acquisition, as this would require retaining historical state longer than necessary for certificate validation.%
\footnote{We do not expect certificate validation to become the bottleneck of syncing, as $\perasRoundSlots \cdot \asc = \num{4.5}$ and \numrange{4}{5} blocks empirically take longer to validate than a certificate.
  Also, we note that certificates can be trivially validated in parallel.}
To accomplish this, we propose to temporarily stop selecting blocks if doing so would cause the immutable tip of the chain to advance while certificate acquisition is lagging behind.
This is a mechanism very similar to the \emph{Limit on Eagerness} in the Cardano Genesis implementation \parencite{genesis-implementation-documentation}, which avoids permanently committing to any chain if a disagreement between competing header chains is yet to be resolved by a density comparison.

%%% Local Variables:
%%% mode: latex
%%% TeX-engine: xetex
%%% TeX-master: "../peras-design"
%%% End:
